---
title: "Correlation Analysis"
author: "ATIQ UR REHMAN"
date: "Date: 2/18/2021"
output:
  beamer_presentation:
    theme: "AnnArbor"
    colortheme: "dolphin"
    fonttheme: "structurebold"
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## INTRODUCTION

- Correlation analysis measures the relationship between the two variables.

- Correlation analysis shows us how to determine both the nature and strength of relationship between two variables.

- Three possible results in the correlational study: Positive, Negative and no correlation.

- Example positive correlation: Income and consumption

- Example of negative correlation: Cold weather and air conditioning cost

- Example of no correlation: price of clothes and car speed.

## Types of Correlation

- Perfect positve correlation

- perfect negative correlation 

- strong/weak positive correlation

- strong/weak negative correlation

- absolute no correlation 

## ...

- A correlation of +1 indicates a perfect positive  correlation.

- A correlation of -1 indicates a perfect negative  correlation.

- Correlation value close to 1 shows strong and close to 0 indicates weak correlation.

- Also correlation value close to -1 shows strong and close to 0 indicates weak correlation.

- A zero correlation indicates that there is no  relationship between the variables.

## Some diagrams of correlatons

```{r,fig.width=1.5, fig.height=1.5, figures-side, fig.show="hold", out.width="50%"}
library(ggplot2)

them <- theme(axis.text.x=element_blank(),
              axis.text.y=element_blank(), 
              axis.ticks.x = element_blank(),
              axis.ticks.y = element_blank())

a <- rnorm(100)
b <- a
d <- 1+rnorm(100)
e <- 0.9*a+rnorm(100)
f <- -0.9*a+rnorm(100)
g <- 0.5*a+rnorm(100)
df <- data.frame(a= a, 
                 b= b, 
                 c = -b,
                 d=d,
                 e=e,
                 f= f,
                 g= g)
par(mfrow=c(2,2))#,mar = c(4, 4, .1, .1)

ggplot(data=df,aes(a,b)) +
  geom_point() +
  theme_dark() +
  them 
  
  
ggplot(data=df,aes(a,c)) +
  geom_point() +
  theme_dark() +
  them 

ggplot(data=df,aes(a,d)) +
  geom_point() +
  theme_dark() +
  them 

ggplot(data=df,aes(a,e)) +
  geom_point() +
  theme_dark() +
  them 

ggplot(data=df,aes(a,f)) +
  geom_point() +
  theme_dark() +
  them 

ggplot(data=df,aes(a,g)) +
  geom_point() +
  theme_dark() +
  them 
```

## Method of studying correlation 

- Scatter Diagram

- Karl Pearson' method of correlation

## Scatter diagram

- Scatter diagram can only tell us about the nature of correlation.

- We can identify the direction of the relation visually.

- We cannot determine the magnitude of correlation in scatter diagram.

## Karl Pearson method of correlation

- It is denoted by "r".

- The formula of r is

$$r= \frac{cov(x,y)}{\sqrt{Var(x)Var(y)}}, $$
where $cov(x,y)$ is the shared variability of the two variables and $Var(x)$, $Var(y)$ are the individual variability each varibale.

- The range of r is $-1 \le r \le +1$



## Real dataset

```{r cars, echo = TRUE}
df_cars <- datasets::cars

str(df_cars)

```

- The correlation between the cars' speed and distance travelled by the cars is
```{r, echo = TRUE, warning=FALSE}
cor(df_cars$speed, df_cars$dist)
```

## The scatter plot of cars' dataset

 
```{r, warning=FALSE,fig.width=2.8, fig.height=2.8}
ggplot(df_cars, aes(x=df_cars$speed, y=df_cars$dist)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  xlab("Speed") +
  ylab("Distance") +
  theme_dark() +
  them
```

